---
title: "Fuel Economy and Automobile Transmission Type"
author: "Michael M. Clark"
date: "March 11, 2015"
output:
  pdf_document: 
    fig_caption: yes
---

#Executive Summary
This brief report explores the questions, "Is an automatic or manual transmission better for MPG?" and "How different is the MPG between automatic and manual transmission?" To answer these questions, data from the 1974 Motor Trend magazine on 32 automobiles were analyzed to look for trends relating fuel economy (i.e. MPG) to transmission type. A quick look at the data shows that median MPG of manual transmission vehicles is higher than that of automatic transmission vehicles, suggesting that manual transmission is better for fuel economy. However the most important factor appears to be vehicle weight, with lower-weight vehicles having higher fuel economy. Most of the lower-weight vehicles in the Motor Trend data were also manual transmission vehicles.

#The MotorTrend Cars Dataset
```{r echo=FALSE}
data(mtcars)
require(graphics)
df <- mtcars
df$cyl.factor  <- as.factor(df$cyl)
df$am <- as.factor(df$am)
df$vs  <- as.factor(df$vs)
df$gear.factor <- as.factor(df$gear)
df$trans <- "x"
df$trans[which(df$am==0)] <- "Automatic"
df$trans[which(df$am==1)] <- "Manual"
df$trans <- as.factor(df$trans)



par(mfrow=c(2, 3))
plot(df$trans, df$mpg, ylab="miles per gallon")
plot(df$cyl.factor, df$mpg, xlab="# of cylinders", ylab="miles per gallon")
plot(df$gear.factor, df$mpg, xlab="# forward gears", ylab="miles per gallon")

plot(df$wt, df$mpg, xlab="weight, 1000lbs", ylab="miles per gallon", col=mtcars$am+1, pch=15)
lm2 <- lm(mpg ~ wt, data=df)
abline(lm2, lwd=3, lty=2)
legend("topright", pch=15, legend=c("automatic", "manual"), col=c(1, 2))

plot(df$wt, df$mpg, xlab="weight, 1000lbs", ylab="miles per gallon", col=mtcars$cyl/2, pch=15)
lm2 <- lm(mpg ~ wt, data=df)
abline(lm2, lwd=3, lty=2)
legend("topright", pch=15, legend=c("4-cylinder", "6-cylinder", "8-cylinder"), col=c(2, 3, 4))

plot(df$hp, df$mpg, xlab="gross horsepower", ylab="miles per gallon",col=mtcars$am+1, pch=15)
lm3 <- lm(mpg ~ hp, data=df)
abline(lm3, lwd=3, lty=2)
legend("topright", pch=15, legend=c("automatic", "manual"), col=c(1, 2))

plot(df$qsec, df$mpg, xlab="1/4 mile time, sec", ylab="miles per gallon",col=mtcars$am+1, pch=15)
lm4 <- lm(mpg ~ qsec, data=df)
abline(lm4, lwd=3, lty=2)
legend("topright", pch=15, legend=c("automatic", "manual"), col=c(1, 2))


pairs(mtcars, panel = panel.smooth, main = "Motor Trend 1974 Cars Data")

```



#Analysis

```{r echo=FALSE}
lm1 <- lm(mpg ~ trans, data=df)
summary(lm1)$coeff

lm2 <- lm(mpg ~ wt, data=df)
summary(lm2)$coeff

lm3 <- lm(mpg ~ wt + wt*trans, data=df)
summary(lm3)$coeff

lm4 <- lm(mpg ~ wt + cyl.factor + wt*trans, data=df)
summary(lm4)$coeff


plot(lm2)

```

#Conclusions


#Appendix



```{r}

```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

```{r mns-hist, echo=FALSE, fig.width=5, fig.height=3, fig.cap="Sample figure caption"}
maintxt <- c(paste("samples of size =", sample.size, "from the exp. dist."), paste("lambda = ", format(lambda, digits=2), ", theoretical mean = ", format(1/lambda, digits=2)))
xlabel <- "sample mean"
hist(mns, breaks=100, main=maintxt, xlab=xlabel)
abline(v=1/lambda, col="red", lwd=3)
```
