---
title: "Fuel Economy and Automobile Transmission Type"
author: "Michael M. Clark"
date: "March 11, 2015"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---
```{r echo=FALSE}
data(mtcars)
require(graphics)
df <- mtcars
df$cyl.factor  <- as.factor(df$cyl)
df$am <- as.factor(df$am)
df$vs  <- as.factor(df$vs)
df$gear.factor <- as.factor(df$gear)
df$trans <- "x"
df$trans[which(df$am==0)] <- "Automatic"
df$trans[which(df$am==1)] <- "Manual"
df$trans <- as.factor(df$trans)
mpg.mt.mean <- mean(df$mpg[which(df$am==1)])
mpg.mt.median <- median(df$mpg[which(df$am==1)])
mpg.at.mean <- mean(df$mpg[which(df$am==0)])
mpg.at.median <- median(df$mpg[which(df$am==0)])
```
#Executive Summary
This brief report explores the questions, "Is an automatic or manual transmission better for MPG?" and "How different is the MPG between automatic and manual transmission?" To answer these questions, data from the 1974 Motor Trend magazine on 32 automobiles were analyzed to look for trends relating fuel economy (i.e. MPG) to transmission type. The mean MPG of manual transmission vehicles (`r round(mpg.mt.mean, digits=3)` mpg) is higher than that of automatic transmission vehicles (`r round(mpg.at.mean, digits=3)` mpg), suggesting that manual transmission is better for fuel economy. However, the most important factor appears to be vehicle weight, with lower-weight vehicles having higher fuel economy. In fact almost all of the lower-weight (<3,000lbs) vehicles in the Motor Trend data were manual transmission vehicles, and almost all of the higher-weight vehicles (>3,000lbs) were automatic transmission vehicles. Recognizing the large effect of vehicle weight on MPG and the insufficient distribution of transmission types across the full range of vehicle weight in the data, we cannot with confidence determine whether tranmission type itself affects MPG. Instead we propose a model for MPG that accounts for the effects of vehicle weight and number of cylinders:

MPG = beta0 + beta1*wt + beta2*

#The MotorTrend Cars Dataset



#Appendix - Figures and Code
```{r exploratory, echo=FALSE, fig.width=6, fig.height=6, fig.cap="Data from the 1974 Motor Trend magazine. A) a boxplot of fuel economy by transmission type indicates that manual transmission vehicles on average have better fuel economy than automatic transmission vehicles. B) a boxplot of fuel economy by # of cylinders indicates that cars with fewer cylinders have better fuel economy. C) a plot of fuel economy by # of forward gears shows a non-monotonic trend. D) a plot of fuel economy vs. weight shows a clear trend of decreasing fuel economy with increasing weight in both manual transmission cars and automatic transmission cars. The average weight of manual transmission cars is lower the average weight of automatic transmission cars. E) a plot of fuel economy vs. weight showing decreasing fuel economy with increasing weight for cars with 4, 6, and 8 cylinder engines. F) a plot of fuel economy vs. horsepower shows decreasing fuel economy with increasing horsepower for both manual and automatic transmission vehicles."}
par(mfrow=c(2, 3))
plot(df$trans, df$mpg, ylab="miles per gallon", main="A")
plot(df$cyl.factor, df$mpg, xlab="# of cylinders", ylab="miles per gallon", main="B")
plot(df$gear.factor, df$mpg, xlab="# forward gears", ylab="miles per gallon", main="C")

plot(df$wt, df$mpg, xlab="weight, 1000lbs", ylab="miles per gallon", col=mtcars$am+1, pch=15, main="D")
#lm2 <- lm(mpg ~ wt, data=df)
#abline(lm2, lwd=3, lty=2)
abline(v=mean(df$wt[which(df$am==1)]), col="red", lty=3)
abline(v=mean(df$wt[which(df$am==0)]), col="black", lty=3)
legend("topright", pch=c(15, 15, -1, -1), legend=c("automatic", "manual", "avg wt", "linear fit"), col=c(1, 2, 1, 1), lty=c(0, 0, 3, 1), bg="white")
lm5 <- lm(mpg ~ wt + trans + wt*trans, data=df)
abline(c(lm5$coef[1], lm5$coef[2]), col="black", lty=1, lwd=1)
abline(c(lm5$coef[1]+lm5$coef[3], lm5$coef[2]+lm5$coef[4]), col="red", lty=1, lwd=1)

plot(df$wt, df$mpg, xlab="weight, 1000lbs", ylab="miles per gallon", col=mtcars$cyl/2, pch=15, main="E")
lm2 <- lm(mpg ~ wt, data=df)
abline(lm2, lwd=3, lty=2)
legend("topright", pch=15, legend=c("4-cylinder", "6-cylinder", "8-cylinder"), col=c(2, 3, 4))

plot(df$hp, df$mpg, xlab="gross horsepower", ylab="miles per gallon",col=mtcars$am+1, pch=15, main="F")
lm3 <- lm(mpg ~ hp, data=df)
abline(lm3, lwd=3, lty=2)
legend("topright", pch=15, legend=c("automatic", "manual"), col=c(1, 2))

#plot(df$qsec, df$mpg, xlab="1/4 mile time, sec", ylab="miles per gallon",col=mtcars$am+1, pch=15)
#lm4 <- lm(mpg ~ qsec, data=df)
#abline(lm4, lwd=3, lty=2)
#legend("topright", pch=15, legend=c("automatic", "manual"), col=c(1, 2))


#pairs(mtcars, panel = panel.smooth, main = "Motor Trend 1974 Cars Data")

```



#Analysis

```{r echo=FALSE}
lm1 <- lm(mpg ~ trans, data=df)
summary(lm1)$coeff

lm2 <- lm(mpg ~ wt, data=df)
summary(lm2)$coeff

lm3 <- lm(mpg ~ wt + wt*trans, data=df)
summary(lm3)$coeff

lm4 <- lm(mpg ~ wt + cyl.factor + wt*trans, data=df)
summary(lm4)$coeff

lm5 <- lm(mpg ~ wt + trans + wt*trans, data=df)
summary(lm5)$coeff

anova(lm2, lm5)

plot(lm2)

lm6 <- lm(mpg ~ wt + wt*cyl.factor, df)
summary(lm6)

lm7 <- lm(mpg ~ wt + cyl.factor, df)
summary(lm7)
anova(lm2, lm7, lm6)

plot(lm7, which=1)
plot(lm7, which=2)
plot(lm7, which=3)
plot(lm7, which=4)
plot(lm7, which=5)
plot(lm7, which=6)

shapiro.test(lm7$residuals)


```

#Conclusions


#Appendix




